Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-01-04 Ben_Manuscripts/structure_paper/Draft.tex
Removed passive voice unless it made sense to keep it

  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control pore architecture at the molecular scale,
  thereby permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. Most membrane-based aqueous separations of
  small molecules can be achieved using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

@@

  %MRS4: are all commercial RO membranes just like this? This seems
  %awfully specific. 
  %BJC3: Yea, pretty much all commercial membranes are TFC polyamide
  % but I edited it to make it less specific.

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically porous membranes made by a phase-inversion
  process.  The most widely used phase-inversion process is immersion
  precipitation, during which a polymer dissolved in a solvent is submerged in a
  non-solvent. A solid, porous polymer membrane is all that remains once all
  solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
  size, which hurts membrane selectivity. A second technique used to create NF
  membranes is call track-etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
  however, the membranes have a low porosity and subsequently low permeability.

@@

  Research into polymerized LLC membranes has been revived in recent years. During early
  stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  the mesophases could be aligned using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that the same result could be obtained using a second technique
  termed soft confinement \cite{feng_thin_2016}.

@@

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure which have been drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the volume of the liquid crystal monomers
  proposes that there are seven monomers in each layer
  \cite{resel_structural_2000}.  A molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulations, as we
  present here, has the best chance of directly answering this question.  We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  Experimentally, their existence is supported by evidence of strong $\pi$-$\pi$
  stacking interactions in the direction perpendicular to the membrane plane.
  $\pi$-$\pi$ stacking will only occur between the aromatic monomer head groups
  which leaves no description of what is happening in the monomer tail region.
  The tails may entangle isotropically while stacking order is maintained among
  headgroups.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self-assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a ``dry'' and ``wet'' system.

@@

  	\item The first is located at $q_z$ = 1.7 \angstrom$^{-1}$, corresponding to
	a real space separation of 3.7 \angstrom~.  The reflection is attributed to
	$\pi$-$\pi$ stacking between aromatic rings in the direction perpendicular to
	the membrane plane, or z-axis \cite{feng_scalable_2014}. For simplicity, this
	reflection will be referred to as R-$\pi$.

	\item A weak intensity line is located at exactly half the $q_z$ value
	of R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponding to a real space
	periodic spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis
	\cite{feng_scalable_2014}, meaning that if the positions of the aromatic rings
	can be traced by a helix, then for each full turn in the helix, one will
	encounter two aromatic rings. For this reason it will be referred to as
	R-helix.

	\item A third major reflection is marked by a low intensity ring
	located at r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to
	4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains \cite{mcintosh_organization_1980}. This reflection will be called
	R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity
	which will be called R-spots. All are located $\sim 37$ degrees from the $q_z$
	axis in their respective quadrants. In many liquid crystal systems this can be
	explained by the tilt angle of the alkane chains with respect to the membrane
	plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane which can be related to the distance between pores.
	The feature, which will be called R-pores, is characterized by dots along $q_z$
	= 0. The spacing between dots is indicative of the hexagonal symmetry of the
	packed pores. The same information at higher resolution is obtained using SAXS (Fig.~\ref{fig:SAXS}).

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one.
%BJC3: Got it, will do once I have more time to focus on figures 
	\caption{(a) (Reproduced from \cite{feng_thin_2016}) The repeat spacing
		in the 1D small angle X-ray scattering pattern is characteristic of hexagonal
		packing. The leading peak represents the distance between the
		d\textsubscript{100} planes. Using this distance, we know that the distance
		between pore centers is 4.12 nm. (b) 2D wide angle X-ray scattering gives
		details about repeating features less than 1 nanometer apart. Experimentalists
		have justified each of the 5 major reflections present as follows: (I) Aromatic
		head group $\pi-\pi$ stack 3.7 \AA~apart.  (II) Monomers arrange vertically in
		a 2\textsubscript{1} helix. (III) Alkane chain tails pack 4.5 \AA~apart. (IV)
		Monomer tails tilt with respect to the membrane plane. (V) As derived from
		SAXS, pores pack hexagonally and are spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  Liquid crystal monomers were parameterized using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. Atomic charges were assigned using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. All molecular dynamics simulations were run using Gromacs 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.
  An ensemble of characteristic, low-energy vacuum monomer
  configurations were constructed by applying a simulated annealing process to a
  parameterized monomer. Monomers were cooled from 1000K to 50K over 10
  nanoseconds.  A low energy configuration was randomly pulled from the
  trajectory and charges were reassigned using \texttt{molcharge}.  Using the new
  charges, the monomer system was annealed again and a random monomer
  configuration was pulled from the trajectory to be used for full system
  construction (Figure~\ref{fig:python}a).  See supplemental information for
  further detail.

@@

%MRS4: check grammar additions in this paragraph.
  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system.  Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system. To bypass the slow
  self-assembly process, Python scripts are used to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) A single monomer was parameterized and annealed to produce a low energy
		configuration. (b) Monomers are rotated and assembled into layers with 
		hydrophilic centers. (c) Twenty layers are stacked on top of each other to create
		a pore. (d) Pores are duplicated and placed into a monoclinic unit cell}\label{fig:python}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
%BJC2: When I first set up the system, there was a gap and a small number of layers 
% appeared to cause the phase to disassociate. However, I can get a pretty small
% number of layers in the infinite pore system, but the diffraction is low resolution.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure when a
  reasonable value is chosen (See Supplemental Information). The pore radius is
  chosen to be 0.6 nm in our initial configurations because the pore size is
  estimated to be $\sim$1.2 nm \cite{zhou_supported_2005}. The initial distance
  between pores also has little effect on the the equilibrated structure (See
  supplemental information). However, one should not start them too close or
  there will be high energy repulsions during early equilibration. We chose an
  initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm.  The distance between layers, the rotation of the layers with
  respect to adjacent layers, and the number of monomers per layer do influence
  the equilibrium structure and require further justification for their choices.
  We rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane.  We explore two different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings are stacked 3.7
  \angstrom~apart in the z-direction. A second system is explored with an initial
  layer spacing of 5 \angstrom. A third system with an initial layer spacing of
  10 \angstrom~was briefly explored. When layers are spaced out too far, they
  will collapse on each other while simultaneously slipping in between layers of
  adjacent pores, which leads to an artificially thick membrane with pores spaced
  closely together.  Further details of simulations with large layer spacing are
  in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002} (Fig. \ref{fig:sandwiched} and \ref{fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\sim$5 \angstrom~
  equilibrium stacking distance \cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail. Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

  The number of monomers in each layer is unknown, as stated in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. Systems were built in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are reduced in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. The resulting
  unrestrained structure is allowed to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

  In order to create a wet system, we solvated an initial configuration with
  water using \texttt{gmx solvate}. All water molecules placed outside the pore
  region are removed. Waters inside the pore region are randomly removed until
  the desired concentration of water in the pores is reached. The remainder of
  the equilibration follows the same procedure as the dry system.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on a
  distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system \cite{young_introduction_2011}.
  Direction of attack was not considered because the resultant mixture is
  racemic.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. Pore centers are located by averaging the coordinates of sodium
  ions in their respective pores. Pore spacing statistics were generated 
  using the bootstrapping technique (See Supplemental Information).

@@

  The colorbars on all diffraction patterns are normalized relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions.  Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.

>>>>

  Nanostructured porous membrane materials have become increasingly popular for
  aqueous separations applications such as desalination and biorefinement because
  they offer the ability to control pore architecture at the molecular scale,
  thereby permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. One can acheive most membrane-based aqueous separations of
  small molecules using reverse osmosis (RO) or nanofiltration
  (NF) \cite{van_der_bruggen_review_2003}. While RO and NF have seen many
  advances in the past few decades, they are far from perfect separation
  technologies.

@@

@@

%  solution.  Through the use of a non-solvent, a polymer-rich and polymer-poor
%  phase is formed. A membrane is created as the polymer-poor phase forms pores in
%  the polymer-rich phase. 
%  During phase inversion, a thin polymer film is precipitated out of a polymer
  NF membranes are typically porous membranes made by a phase-inversion
  process.  The most widely used phase-inversion process is immersion
  precipitation, during which one submerges a polymer, dissolved in a solvent, in
  a non-solvent. A solid, porous polymer membrane is all that remains once all
  solvent has been removed by non-solvent exchange
  \cite{smolders_microstructures_1992}.  The resultant pores are polydisperse in
  size, which hurts membrane selectivity. A second technique used to create NF
  membranes is call track-etching in which a polymer film is bombarded with
  charged particles, then chemically etched to create pores
  \cite{apel_track_2001}. The pores are uniform, which benefits selectivity;
  however, the membranes have a low porosity and subsequently low permeability.

@@

  Research into polymerized LLC membranes has been revived in recent years.
  During early stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  one can align the mesophases using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}.

@@

  In this study, we build a significantly more realistic atomistic model of LLC
  membranes than, to our knowledge, has ever previously been done, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the model using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small angle X-ray scattering (SAXS)
  and wide angle X-ray scattering (WAXS) experiments as well as in matching ionic
  conductivity measurements \cite{feng_thin_2016}.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the volume of the liquid crystal monomers
  proposes that each layer consists of seven monomers
  \cite{resel_structural_2000}.  A molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulations, as we
  present here, has the best chance of directly answering this question.  We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  Strong $\pi$-$\pi$ stacking interactions in the direction perpendicular to
  the membrane plane support the existence of layers experimentally.  $\pi$-$\pi$
  stacking will only occur between the aromatic monomer head groups which leaves
  no description of what is happening in the monomer tail region.  The tails may
  entangle isotropically while maintaing stacking order among headgroups.

@@

  While past work describes the Col\textsubscript{h} phase as dry
  ~\cite{feng_scalable_2014}, it has been suggested by experimentalists, in
  unpublished communications, that it is likely that neat monomer leaches small
  amounts of ambient water.  Experimentally, achieving a hexagonal phase with a
  completely dry system has proven difficult. If neat monomer sits in ambient
  conditions, its color turns from transparent to slightly opaque and a hexagonal
  phase forms.  Although we will not explore whether water is necessary for
  self-assembly, we hypothesize that the hydrogen bonding network formed by the
  water may play a role in structuring the pores and holding together the
  hexagonal phase. We can use simulated X-ray diffraction patterns to see if
  there is any meaningful structural difference between a ``dry'' and ``wet''
  system.

@@

	\item The location of the first is at $q_z$ = 1.7 \angstrom$^{-1}$,
	corresponding to a real space separation of 3.7 \angstrom~. Previous
	work~\ref{feng_scalable_2014} attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

	\item A weak intensity line, located at exactly half the $q_z$ value of
	R-$\pi$ ($q_z$ = 0.85 \angstrom$^{-1}$), corresponds to a real space periodic
	spacing of 7.4 \angstrom~. This reflection has been interpreted as
	2\textsubscript{1} helical ordering of aromatic rings along the z axis
	\cite{feng_scalable_2014}, meaning that if one traces the positions of the
	aromatic rings with a helical curve, then for each full turn in the helix, one
	will encounter two aromatic rings. For this reason we will refer to reflection
	as R-helix.

	\item A low intensity ring located at r = 1.4 \angstrom$^{-1}$ marks
	the third major reflection of interest. The real space separation corresponds
	to 4.5 \angstrom~ which is characteristic of the average spacing between packed
	alkane chains \cite{mcintosh_organization_1980}. We will call this reflection
	R-alkanes.

	\item Within R-alkanes, are four spots of higher relative intensity.
	Accordingly, we name these reflection R-spots. Thin location of all spots is $\sim 37$
	degrees from the $q_z$ axis in their respective quadrants. In many liquid
	crystal systems one can explain the spots by the tilt angle of the alkane chains
	with respect to the membrane plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane. This plane is geometrically related the distance
	between pores.  The feature, which we named R-pores, is characterized by dots
	along the equatorial axis defined when $q_z$ = 0. The spacing between dots is
	indicative of the hexagonal symmetry of the packed pores. We observe the same
	information at higher resolution using SAXS (Fig.~\ref{fig:SAXS}).

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS3: would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
%BJC2: I will annotate it with the names R-pi, R-alkanes etc. Is that what you mean?
%MRS4: that works fine.  We just need a way to connect what is in the figure with the names used in the text.  Ideally there would not be two sets of labels, just one.
%BJC3: Got it, will do once I have more time to focus on figures 
	\caption{(a) (Reproduced from \cite{feng_thin_2016}) The repeat spacing
		in the 1D small angle X-ray scattering pattern is characteristic of hexagonal
		packing. The leading peak represents the distance between the
		d\textsubscript{100} planes. Using this distance, we know that the distance
		between pore centers is 4.12 nm. (b) 2D wide angle X-ray scattering gives
		details about repeating features less than 1 nanometer apart. Experimentalists
		have justified each of the 5 major reflections present as follows: (I) Aromatic
		head group $\pi-\pi$ stack 3.7 \AA~apart.  (II) Monomers arrange vertically in
		a 2\textsubscript{1} helix. (III) Alkane chain tails pack 4.5 \AA~apart. (IV)
		Monomer tails tilt with respect to the membrane plane. (V) As derived from
		SAXS, the pores are spaced 4.12 nm apart and pack hexagonally}
	\label{fig:SWAXS}
 \end{figure}

@@

  We parameterized the liquid crystal monomers using the Generalized AMBER
  Forcefield \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We assigned atomic charges using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. We ran all molecular dynamics simulations using Gromacs 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

  We constructed an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000K to 50K over 10
  nanoseconds. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration the trajectory which we used for full system
  construction (Figure~\ref{fig:python}a). See supplemental information for
  further detail.

@@

%MRS4: check grammar additions in this paragraph.
  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system.  Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}.  Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system. To bypass the slow
  self-assembly process, we use Python scripts to assemble monomers into a
  structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:python}).

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) Python scripts rotate and assemble monomers into layers with 
		hydrophilic centers. (c) We chose to stack twenty layers on top of each other to create
		pores. (d) The pores are duplicated and placed into a monoclinic unit cell}\label{fig:python}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number of layers
  in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
%BJC2: When I first set up the system, there was a gap and a small number of layers 
% appeared to cause the phase to disassociate. However, I can get a pretty small
% number of layers in the infinite pore system, but the diffraction is low resolution.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. The starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one choses
  a reasonable value (See Supplemental Information). The pore radius is chosen to
  be 0.6 nm in our initial configurations because the pore size is estimated to
  be $\sim$1.2 nm \cite{zhou_supported_2005}. The initial distance between pores
  also has little effect on the the equilibrated structure (See supplemental
  information). However, one should not start them too close or there will be
  high energy repulsions during early equilibration. We chose an initial pore
  spacing of 4.5 nm, $\sim$10\% larger than the experimental value of 4.12 nm.
  The distance between layers, the rotation of the layers with respect to
  adjacent layers, and the number of monomers per layer do influence the
  equilibrium structure and require further justification for their choices.  We
  rely on experimental data to inform them.

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane. We explore two different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial layer spacing of 5
  \AA. We briefly explored a third system with an initial layer spacing of 10
  \AA. If we initially space layers too far apart, they will collapse on each
  other while simultaneously slipping in between layers of adjacent pores, which
  leads to an artificially thick membrane with pores spaced closely together.
  Further details of simulations with large layer spacing are in the supplemental
  information.

  We chose the relative interlayer orientation based on clues from diffraction
  data as well as the various known stacking modes of benzene and substituted
  benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002} (Fig. \ref{fig:sandwiched} and
  \ref{fig:pd,fig:tshaped}).  We ruled out the T-shaped configuration because its
  $\sim$5 \angstrom~ equilibrium stacking distance
  \cite{sinnokrot_estimates_2002} is inconsistent with R-$\pi$. It is also
  infeasible for the monomers to orient in the T-shaped conformation because of
  the bulky tail groups. We will explore the system's preference towards the
  sandwiched vs. parallel displaced stacking modes in some detail. Both have
  reported stacking distances near the R-$\pi$ value of 3.7 \angstrom. Headgroups
  in our sandwiched initial configuration stack directly on top of each other
  while headgroups in the parallel displaced initial configuration stack
  with an offset of $180/nmon$ degrees where $nmon$ equals the number of monomers
  per layer.

  The number of monomers in each layer is unknown, as stated in
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of monomers per layer. We built systems in the offset and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

  In order to create a wet system, we solvated an initial configuration with
  water using \texttt{gmx solvate}. We remove all water molecules placed outside
  the pore region. Then we randomly remove water molecules inside the pore region
  until the pores reach the desired concentration of water. The remainder of the
  equilibration follows the same procedure as the dry system.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that one can apply to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. We
  developed the algorithm based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, we simulated the initiator as hydrogen and made it
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. We carry out the crosslinking procedure
  iteratively. During each iteration, the algoritm selects eligible bonding
  carbon atoms based on a distance cut-off. The topology is updated with new
  bonds and dummy hydrogen atoms are changed to appropriate hydrogen types.
  Head-to-tail addition was the only propagation mode considered due to its
  dominance in the real system \cite{young_introduction_2011}. We did not
  consider direction of attack because the resultant mixture is racemic.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of sodium
  ions in their respective pores. We generated pore spacing statistics 
  using the bootstrapping technique (See Supplemental Information).

@@

  We normalized the colorbars on all diffraction patterns relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions.  Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.
